--
title: "Data Wrangling Ex 1_refine1"
author: "H Karen Tam"
date: "October 20, 2017"
output: html_document
---
### Data Wrangling Exercise 1: Basic Data Manipulation

#### Load Packages and Read csv File
```{r}

# Load package

suppressMessages(library(dplyr))
suppressMessages(library(tidyr))

# Read the csv file

refine_clean <- read.csv("refine_original.csv")

```

#### Clean Up Brand Names

```{r}

## Clean up brand names in the "company"" column

# Transform values in the column: philips, akzo, van houten and unilever (all lowercase)

refine_clean <- refine_clean %>% 
mutate(company = tolower(company)) %>% 
mutate(company = gsub("phillips|phllips|phillps|fillips|phlips", "philips", company)) %>% 
mutate(company = gsub("unilver", "unilever", company)) %>% 
mutate(company = gsub("akz0|ak zo", "akzo", company))

```

#### Separate Product Code and Number

```{r}

## Separate product code and number

# Add two new columns called product_code and product_number, containing the product code and number respectively.

refine_clean <- refine_clean %>% 
  separate(Product.code...number, c("product_code", "product_number"), sep = "-")

```

#### Add Product Categories

```{r}
## Add product categories

# Add a column with the product category for each record. p = Smartphone; v = TV; x = Laptop; and q = Tablet.

refine_clean$product_category <- refine_clean$product_code

# Change values from p, v, x, or q to Smartphone, TV, Laptop, or Tablet

refine_clean <- refine_clean %>% 
mutate(product_category = gsub("p", "smartphone", product_category)) %>% 
mutate(product_category = gsub("v", "tv", product_category)) %>% 
mutate(product_category = gsub("x", "laptop", product_category)) %>% 
mutate(product_category = gsub("q", "tablet", product_category))
```

#### Add Full Address for Geocoding

```{r}
## Add full address for geocoding

# Create a new column full_address that concatenates the three address fields (address, city, country), separated by commas.

refine_clean <- refine_clean %>% 
mutate(full_address = paste(address, city, country, sep = ", " ))
```

#### Create Dummy Variables For Company and Product Category

```{r}
## Create dummy variables for company and product category

# Create dummy binary variables: 
# Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever.
# Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.

refine_clean <- refine_clean %>% 
mutate(company_philips = as.numeric(as.logical(company == "philips"))) %>% 
mutate(company_akzo = as.numeric(as.logical(company == "akzo"))) %>% 
mutate(company_van_houten = as.numeric(as.logical(company == "van houten"))) %>% 
mutate(company_unilever = as.numeric(as.logical(company == "unilever"))) %>% 
mutate(product_smartphone = as.numeric(as.logical(product_category == "smartphone"))) %>% 
mutate(product_tv = as.numeric(as.logical(product_category == "tv"))) %>% 
mutate(product_laptop = as.numeric(as.logical(product_category == "laptop"))) %>% 
mutate(product_tablet = as.numeric(as.logical(product_category == "tablet")))
```

#### Reorder Columns by Position

```{r}
## Reorder columns by position

refine_clean <- refine_clean [, c(1, 8, 2, 3, 4, 5, 6, 9, 7, 10, 11, 12, 13, 14, 15, 16, 17)]
  
```  

#### Save Cleaned Dataframe as .csv

```{r}
## Save the cleaned file as .csv

write.csv(refine_clean, "refine_clean.csv", row.names = FALSE)

```

